---
title: "Current Users"
author: ""
date: ""
output: html_document
---

```{r, message = FALSE, echo = FALSE,results='hide', warning=FALSE}
library(tidyverse)
library(here)
library(grid) #for Grobs
library(scales) #pretty breaks

# Inherit `resultsTidy`
knitr::knit_child(here("TidyData.Rmd"))
# Import functions to stylize plots
source(here("resources/scripts/shared_functions.R"))
```

<br>

# Length of Use of the AnVIL

```{r, message = FALSE, echo = FALSE}
timeUsePlot <- resultsTidy %>%
  group_by(LengthOfUse) %>%
  summarize(count = n()) %>%
  drop_na() %>%
  ggplot(aes(x = LengthOfUse,
             y = count,
             fill = "#25445A")) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count, group = LengthOfUse),
                  vjust = -1, size=2) +
  ggtitle("How long have you been using the AnVIL?")

stylize_bar(timeUsePlot, usertypeColor = FALSE, singleColor = TRUE) +
    xlab("Years of Use") +
    ylab("Count") +
    theme(legend.position = "none")
```

## Takeaway

We observe a fairly even sampling of the current users with regards to the number of years they've been using the AnVIL

<br>

# Foreseeable Computational Needs

```{r, message=FALSE, echo=FALSE, warning=FALSE}
compNeedsPlot <- resultsTidy %>%
  separate(NeededResources,
           c("whichResourceA", "whichResourceB", "whichResourceC", "whichResourceD"),
           sep = ", ", fill = "right") %>%
  pivot_longer(starts_with("whichResource"), values_to = "ResourceDescription") %>%
  group_by(ResourceDescription) %>%
  summarize(count = n()) %>%
  drop_na() %>%
  ggplot(aes(x = count,
             y = reorder(ResourceDescription, count),
             fill = "#25445A")) +
  geom_text(aes(label = count, group = ResourceDescription),
                  hjust = -1, size=2) +
  geom_bar(stat = "identity") +
  ggtitle("What computational and storage resources do you foresee\nneeding in the next 12 months?")

  stylize_bar(compNeedsPlot, usertypeColor = FALSE, singleColor = TRUE) +
    theme(legend.position = "none")
```

## Takeaway

Of the `r nrow(resultsTidy %>% filter(UserType == "Current User"))` current users, all `r 50 - sum(is.na(resultsTidy$NeededResources))` provided an answer to this question. The most common response here is needing large amounts of storage.

## Potential Follow-ups

  - What do we see in other responses from the "don't know" responses?

<br>

# Recommendation Likelihood

```{r, message = FALSE, echo = FALSE}
recLikePlot <- resultsTidy %>%
  group_by(RecommendationLikelihood) %>%
  summarize(count = n()) %>%
  drop_na() %>% #not asked to everyone
  ggplot(aes(x = RecommendationLikelihood,
             y = count,
             fill = as.factor(RecommendationLikelihood))) +
  geom_bar(stat="identity") +
  ggtitle("How likely are you to recommend the AnVIL to a colleague?") +
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(1,1,1,1.1, "cm")) +
  annotation_custom(textGrob("Extremely\nlikely", gp=gpar(fontsize=8, fontface = "bold")),xmin=5,xmax=5,ymin=-1.25,ymax=-1.25) +
  annotation_custom(textGrob("Not at all\nlikely", gp=gpar(fontsize=8, fontface= "bold")),xmin=1,xmax=1,ymin=-1.25,ymax=-1.25) +
  scale_y_continuous(breaks= pretty_breaks()) +
  geom_text(aes(label = count, group = RecommendationLikelihood),
                  vjust = -1, size=2)

  stylize_bar(recLikePlot, usertypeColor = FALSE, sequentialColor = TRUE) +
    xlab("Recommendation likelihood") +
    ylab("Count") +
    theme(legend.position = "none")
```

## Takeaway

There's a fairly bimodal distribution here with users either extremely likely or only moderately likely to recommend the AnVIL.

## Potential Follow-ups

  - What do we see in other responses from those who are extremely likely to recommend the AnVIL to a colleague? e.g., Are they part of a consortium?
  - What do we see in other responses from those who are only moderately or not at all likely to recommend the AnVIL?
