---
title: "Demographics"
author: ""
date: ""
output: html_document
---

```{r, message = FALSE, echo = FALSE,results='hide', warning=FALSE}
library(tidyverse)
library(here)
library(grid) #for Grobs
library(scales) #pretty breaks
library(kableExtra)

# Inherit `resultsTidy`
knitr::knit_child(here("TidyData.Rmd"))
# Import functions to stylize plots
source(here("resources/scripts/shared_functions.R"))
```

<br>

# Highest Degree

```{r, message=FALSE, echo=FALSE}
degreePlot <- resultsTidy %>%  
  group_by(FurtherSimplifiedDegrees, UserType) %>%
  summarize(n = n()) %>%
  ggplot(aes(y = reorder(FurtherSimplifiedDegrees, n, sum),
             x = n,
             fill = UserType
             )) +
      geom_bar(position = "stack", stat="identity") +
      geom_text(
                  aes(label = after_stat(x), group = FurtherSimplifiedDegrees),
                  stat = 'summary', fun = sum, hjust = -1, size=2
                ) +
      coord_cartesian(clip = "off") +
      ggtitle("What is the highest degree you have attained?")

stylize_bar(degreePlot)
```

## Takeaway

Most of the respondents have a PhD or are currently working on a PhD, though a range of career stages are represented.

<br>

# Kind of Work

```{r, message = FALSE, echo = FALSE}
dfForPlotKOW <- resultsTidy %>%
  separate(KindOfWork,
           c("whichWorkA", "whichWorkB", "whichWorkC", "whichWorkD", "whichWorkE", "whichWorkF", "whichWorkG", "whichWorkH", "whichWorkI", "whichWorkJ"),
           sep=", ", fill="right") %>%
  pivot_longer(starts_with("whichWork"), values_to = "whichWorkDescription") %>%
  select(Timestamp, UserType, whichWorkDescription) %>%
  mutate(whichWorkDescription =
           recode(whichWorkDescription, 
                  "computational education" = "Computational education", 
                  "Program administration," = "Program administration"),
         whichWorkDescription = factor(whichWorkDescription),
         Timestamp = factor(Timestamp)
         ) %>%
  drop_na()

factorLevel <-  as.data.frame(table(dfForPlotKOW$whichWorkDescription)) %>% arrange(-Freq) %>% select(Var1) %>% unlist() %>% unname() %>% rev()

ggplot(dfForPlotKOW, 
       aes(x = Timestamp,
           y = factor(whichWorkDescription, levels = factorLevel),
           fill = whichWorkDescription
           )) +
  geom_tile() +
  theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "None") +
  ylab("") +
  ggtitle("What kind of work do you do?") +
  xlab("Respondent") +
  facet_wrap(~UserType)
```

## Takeaway

Only a few responses showed project management, leadership, or administration as their only kind of work. This increases our confidence that this won't confound later questions asking about usage of datasets or tools.

## Potential Follow-ups

  - Use this together with other info to try to cluster respondents/users into personas.

<br>

# Institutional Affiliation

```{r, message=FALSE, echo = FALSE}
instPlot <- resultsTidy %>%
  mutate(FurtherSimplifiedInstitutionalType = 
           factor(FurtherSimplifiedInstitutionalType, 
                  levels = c("Industry & Other", "Education Focused", "Research Intensive"))) %>%
  group_by(UserType, FurtherSimplifiedInstitutionalType) %>%
  summarize(InstitutionalCount = n()) %>%
  ggplot(aes(
    y = FurtherSimplifiedInstitutionalType,
    x = InstitutionalCount,
    fill = UserType
  )) +
  geom_bar(position = "stack", stat = "identity") +
  geom_text(aes(label = after_stat(x), 
                group = FurtherSimplifiedInstitutionalType),
            stat = 'summary', fun = sum, hjust = -1, size=2
           ) +
  annotation_custom(textGrob("- R1 University     \n- Med Campus      \n- Research Center\n- NIH                     ", gp = gpar(fontsize = 8)), xmin = -8.5, xmax = -8.5, ymin = 2.65, ymax = 2.65) +
  annotation_custom(textGrob("- Industry             \n- International Loc\n- Unknown           ", gp = gpar(fontsize = 8)), xmin = -8.5, xmax = -8.5, ymin = .7, ymax = .7) +
  annotation_custom(textGrob("- R2 University         \n- Community College", gp=gpar(fontsize=8)),xmin=-8.5,xmax=-8.5,ymin=1.75,ymax=1.75) +
  coord_cartesian(clip = "off") +
  ggtitle("What institution are you affiliated with?")

stylize_bar(instPlot)
```

## Takeaway

Most of the individuals using the AnVIL report being affiliated with a research intensive institution. A further breakdown of these counts are below.

## Potential Follow-ups

  - Does industry not being represented in this sample of current users generalize to the larger AnVIL user base? If so, why?

### Related Plots

```{r, message = FALSE, echo = FALSE}
instPlotB <- resultsTidy %>%
  select(c(UserType, InstitutionalType)) %>%
  group_by(UserType, InstitutionalType) %>% 
  summarize(InstitutionalCount = n()) %>%
  ggplot(aes(
    y = reorder(InstitutionalType, InstitutionalCount, sum),
    x = InstitutionalCount,
    fill = UserType
  )) + 
  geom_bar(position = "stack", stat = "identity") +
  geom_text(
    aes(label = after_stat(x), group = InstitutionalType),
    stat = 'summary', fun = sum, hjust = -1, size=2
            ) +
  annotation_custom(textGrob(paste("There are\n", length(unique(resultsTidy$InstitutionalAffiliation))  ,"\nunique institutions"), gp=gpar(fontsize=8, fontface = "bold")),xmin=34,xmax=34,ymin=2.5,ymax=2.5) +
  coord_cartesian(clip = "off") +
  ggtitle("What institution are you affiliated with?")

stylize_bar(instPlotB)
```

<br>

# Consortia Affiliations

```{r, message=FALSE, echo = FALSE}
consortiaTable <- resultsTidy %>%
  mutate(ConsortiaAffiliations = str_replace_all(ConsortiaAffiliations, c(";|&| and"), ",")) %>%
  separate(ConsortiaAffiliations,
           c("whichConsortiumA", "whichConsortiumB", "whichConsortiumC", "whichConsortiumD"),
           sep=", ", fill = "right") %>%
  pivot_longer(starts_with("whichConsortium"), values_to = "whichConsortiumName") %>%
  group_by(whichConsortiumName) %>%
  summarize(count = n()) %>%
  drop_na() %>%
  arrange(count)
```


Of `r nrow(resultsTidy)` responses, `r sum(!is.na(resultsTidy$ConsortiaAffiliations))` provided an affiliation, with `r nrow(consortiaTable)` unique affiliations provided across those responses (respondents could select more than one consortium). The following table shows the most represented consortia.

```{r, message = FALSE, echo = FALSE}
consortia_df <- 
  consortiaTable[which(consortiaTable$count >1),] %>% 
  rename(`consortium` = whichConsortiumName)

kableExtra::kable(consortia_df, table.attr = "style='width:20%;'")
```

